---
title: "755_HW1"
author: "Larri Miller"
date: "due 9/4/2020"
output: pdf_document
---
This HW assignment includes QSS ch. 1 exercise 1.5.1 and QSS ch. 2 exercise 2.8.1.


# 1.5.1 Bias in Self-Reported Turnout
## 1. Load the data into R, check its dimensions, & obtain a data summary.

```{r setup, include=TRUE}

library(tidyverse)
turnout <- read.csv("turnout.csv")
dim(turnout)
summary(turnout)
range(turnout$year)
```
## How many observations are there?
This dataset has `r nrow(turnout)` observations (rows) of `r ncol(turnout)` variables (columns). 

## What is the range of years covered in this data set?
The range of years covered in the dataset is 1980-2008.

## 2. Calculate the turnout rate based on the voting age population (VAP).
```{r}
turnout[is.na(turnout)] = 0 #turning NA values into 0

turnoutVAP <- (turnout$total)/(turnout$VAP + turnout$overseas) * 100
turnoutVAP
```
## Next, calculate the turnout rate using the voting eligible population (VEP).

```{r}
turnoutVEP <- (turnout$total)/ (turnout$VEP + turnout$overseas) * 100
turnoutVEP
```
## What difference do you observe?
The turnout rate calculated using VEP (voting eligible population) is higher than the turnout rate using VAP (voting age population).


## 3. Compute the differences between the VAP and ANES estimates of turnout rate.
```{r}
VAPdif <- abs(turnoutVAP - turnout$ANES)
VAPdif
```
## How big is this difference on average?
```{r}
mean(VAPdif)
```
## What is the range in the differences?
```{r}
range(VAPdif)
```
## Conduct the same comparison for the VEP and ANES estimates of voter turnout. Briefly comment on the results.
```{r}
VEPdif <- abs(turnoutVEP - turnout$ANES)
VEPdif
mean(VEPdif)
range(VEPdif)
```
The difference is smaller for VEP and ANES than it is for VAP and ANES. 


## 4. Compare the VEP turnout rate with the ANES turnout rate separately for presidential elections and midterm elections
First I am looking at the years value in the turnout dataset to see how I should figure out presidential vs. midterm elections.
```{r}
print(turnout$year)
```
In this dataset, the first entry is a presidential election, second is a midterm election, and so on. This means that odd indexes are presidential elections, and even indexes are midterm elections. The dataset skips the year 2006, so the exception to this pattern is that the last value (at i = 14) is actually a presidential election.

### Presidential Elections
```{r}
PresElect <- abs(VEPdif[c(1, 3, 5, 7, 9, 11, 13, 14)] - turnout$ANES[c(1, 3, 5, 7, 9, 11, 13, 14)])
PresElect
mean(PresElect)
range(PresElect)
```
### Midterm Elections
```{r}
midElect <- abs(VEPdif[c(2, 4, 6, 8, 10, 12)] - turnout$ANES[c(2, 4, 6, 8, 10, 12)])
midElect
mean(midElect)
range(midElect)
```
## Does the bias of the ANES estimates vary across election types?
The ANES bias is present in both presidential and midterm election data, but more so in the Presidential; the mean and range in difference values is greater for the Presidential data than the midterm data.

## 5. Divide the data into half by election years such that you subset the data into two periods.
```{r}
period1 <- turnout[1:7, ]
period1
period2 <- turnout[8:14, ]
period2

```
## Calculate the difference between the VEP turnout rate and the ANES turnout rate separately for each year within each period.

### Period 1 (1980-1992)
Calculating VEP turnout rate
```{r}
VEPturnout1 <- (period1$total + period1$osvoters) / (period1$VEP + period1$overseas) * 100
```
Calculating difference
```{r}
period1diff <- abs(VEPturnout1 - period1$ANES)
period1diff
```

### Period 2 (1994-2008)
Calculating VEP turnout rate
```{r}
VEPturnout2 <- (period2$total + period2$osvoters) / (period2$VEP + period2$overseas) * 100
```
Calculating difference
```{r}
period2diff <- abs(VEPturnout2 - period2$ANES)
period2diff
```
## Has the bias of ANES increased over time?
Mean bias for period 1
```{r}
mean(period1diff)
```
Mean bias for period 2
```{r}
mean(period2diff)
```
Yes, the mean value of difference between VEP and ANES is higher for period 2 than period 1, leading to the conclusion that the bias of ANES has increased over time.

## 6. ANES does not interview prisoners and overseas voters. Calculate an adjustment to the 2008 VAP turnout rate.
```{r}
adjVAP <- turnout$VAP[14] - turnout$felons[14] - turnout$noncit[14]

adjVAP2008 <- (turnout$total[14] - turnout$osvoters[14]) / (adjVAP - turnout$overseas[14]) * 100
adjVAP2008
```

## Compare the adjusted VAP turnout wiht the unadjusted VAP, VEP, and the ANES turnout rate.
Unadjusted VAP
``` {r}
turnoutVAP2008 <- (turnout$total[14] + turnout$osvoters[14])/(turnout$VAP[14] + turnout$overseas[14]) * 100
turnoutVAP2008
```

Unadjusted VEP
```{r}
turnoutVEP2008 <- (turnout$total[14] + turnout$osvoters[14])/ (turnout$VEP[14] + turnout$overseas[14]) * 100
turnoutVEP2008
```

ANES
```{r}
turnout$ANES[14]
```
The unadjusted VAP has the lowest turnout rate, with a value of 56. The ANES has the highest turnout rate, at 78. The adjusted VAP and unadjusted VEP fall in the middle of these values, with 64 and 60, respectively.


# 2.8.1 Efficacy of Small Class Size in Early Education
```{r}
star <- read.csv("STAR.csv")
```
## 1. Create a factor variable called kinder in the data frame.
Recode classtype -- following method used in book
```{r}
star$kinder <- NA
star$kinder[star$classtype == "1"] <- "small"
star$kinder[star$classtype == "2"] <- "regular"
star$kinder[star$classtype == "3"] <- "regularWithAid"
star$kinder <- as.factor(star$kinder)

table(star$kinder)
```
Recode race -- using a method learned in DACSS 601
``` {r}
star <- star %>%
  mutate(race = recode(race, `1` = "White", `2` = "Black", `3` = "Asian", `4` = "Hispanic", `5` = "Native American", `6` = "Others"), na.rm=TRUE) 

star$race <- as.factor(star$race)

```

## 2. How does performance on fourth-grade reading and math tests for those students assigned to a small class in kindergarten compare with those assigned to a regular-sized class? Do students in the smaller classes perform better?

Reading tests 
```{r}
tapply(star$g4reading, star$kinder, mean, na.rm = TRUE)
```
Math tests
```{r}
tapply(star$g4math, star$kinder, mean, na.rm = TRUE)
```

## Give a brief substantive interpretation of the results.

In reading classes, students in smaller classes perform slightly better than those in regular classes (723 vs 720). However, in math classes, students in small and regular classes performed about the same (709.19 vs. 709.52). 

## To understand the size of the estimated effects, compare them with the standard deviation of the test scores.
Reading scores standard deviation
```{r}
tapply(star$g4reading, star$kinder, sd, na.rm=TRUE)
```

Math scores standard deviation
```{r}
tapply(star$g4math, star$kinder, sd, na.rm=TRUE)
```

## 3. Look at the entire range of scores of reading and math tests between those students assigned to small classes & those assigned to regular-sized classes.

Reading tests
```{r}
tapply(star$g4reading, star$kinder, summary, na.rm = TRUE)
```
Math tests
```{r}
tapply(star$g4math, star$kinder, summary, na.rm = TRUE)
```
## Does this analysis add anything to the analysis based on mean in the previous question?
All of the differences in scores are very small. For reading, the small classroom had a higher low score than the regular classroom (697.0 vs. 693.0), but the high value was only slightly different (750.0 vs. 749.2). For math, the small classroom had a lower score than the regular classroom (686.8 vs. 688.0) but a greater high score than the regular classroom (736.2 vs. 731.8). Based on these numbers, it seems that smaller classrooms are beneficial in helping kids with lower scores in reading and higher scores in math.

## 4. Some students were in small classes for all four years that the STAR program ran. Others were assigned to small classes for only one year and had either regular-sized classes or regular-sized classes. How many students of each type are in the dataset?
contingency table
```{r}
addmargins(table(star$kinder, star$yearssmall))
addmargins(prop.table(table(star$kinder, star$yearssmall)))
```

## Does participation in more years of small classes make a greater difference in test scores?
Reading scores mean
```{r}
tapply(star$g4reading, star$yearssmall, mean, na.rm = TRUE)
```
Reading scores median
```{r}
tapply(star$g4reading, star$yearssmall, median, na.rm = TRUE)
```

Math scores mean
```{r}
tapply(star$g4math, star$yearssmall, mean, na.rm = TRUE)
```
Reading scores median
```{r}
tapply(star$g4math, star$yearssmall, median, na.rm = TRUE)
```

Overall, the difference between 4 years of small classrooms vs. 0 years of small classrooms is slight; for reading the mean difference is ~ 5 and the median difference is 4 whereas for math it's ~ 2 and 1.

## 5. Examine whether the STAR program reduced achievement gaps across different racial groups. Begin by comparing the average reading and math test scores between white and minority students (i.e., Blacks and Hispanics) among those students who were assigned to regular-sized classes with no aid.

subsetting the data to represent race 
```{r}
white <- subset(star, race == "White")
minority <- subset(star, race == "Black" | race == "Hispanic")
```

Reading scores
```{r}
tapply(white$g4reading, white$kinder=="regular", mean, na.rm = TRUE)
tapply(minority$g4reading, minority$kinder=="regular", mean, na.rm = TRUE)
```
Math score
```{r}
tapply(white$g4math, white$kinder=="regular", mean, na.rm = TRUE)
tapply(minority$g4math, minority$kinder=="regular", mean, na.rm = TRUE)
```
## Give a brief substantive interpretation of the results of your analysis
For both small and regular classes, white children had higher scores than minority children in reading and math.

## 6. Compare high-school graduation rates across students assigned to different class types.
```{r}
addmargins(table(star$hsgrad, star$kinder))
tapply(star$hsgrad, star$kinder, mean, na.rm=TRUE)
```
## Examine whether graduation rates differ depending on the number of years spent in small classes.
``` {r}
addmargins(table(star$hsgrad, star$yearssmall))
tapply(star$hsgrad, star$yearssmall, mean, na.rm = TRUE)
```
## Investigate wheter the STAR program has reduced the racial gap between white and minority student graduation rates.
```{r}
addmargins(table(star$hsgrad, star$race))
tapply(star$hsgrad, star$race, mean, na.rm = TRUE)
```

## Briefly discuss the results.
Students assigned to different class types had nearly identical graduation rates. The mean difference between 4 years of small classrooms vs. 0 years of small classrooms was only ~ 5%. Finally, the difference in mean graduation rate for white students vs. black students (there are no hispanic students in the record) was ~13%.